---
title: "Lab 6"
author: "Alessandra Puig-Santana"
date: "11/2/2021"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)

library(tidyverse)
library(palmerpenguins)
library(broom)
library(equatiomatic)
```

## Ranked based test example
- Mann-Whitney U (non-parametric alternative to an unpaired 2-sample t-test).

H0: Equal ranks (medians) 
HA: Ranks are not equal (medians are not equal)

### Create our mock samples

`set.seed()` to create "pseudorandom, but reproducible samples"

```{r}
set.seed(1414)
gp_1 <- sample.int(20, size = 15, replace = TRUE) # Creating our samples

set.seed(1424)
gp_2 <- sample.int(30, size =15, replace = TRUE) # Creating our samples
```

Tkae a quick look at the distributions using `hist()`
```{r}
hist(gp_1) # exploratory visualization
hist(gp_2) #exploratory visualization
```


Try a t-test so that we can compare the results
```{r}
ttest <- t.test(gp_1, gp_2)

# If these data were paired, I can just add the argument `paired = TRUE` within the t.test() function.

```

This p-value means that IF our samples are drawn from populations with the same mean (i.e., if the null hypothesis is true), there is `r round(ttest$p.value, 2)` probability (chance) of finding sample means *at least as different* as mine by random chance. 

It is pretty likely because it is above 0.05, so we fail to reject the H0. AKA there is no significant difference.

[A more comprehensive explanation of difference between groups... actually means, data spread, effect size].. there is no significant difference (t(df) = t-statistic value, p = ). 

### Man Whitney U rank based test

We use the `wilcox.test()` function to perform a two-sample rank-based test
```{r}
mwu <- wilcox.test(gp_1, gp_2)
```

This p-value means that IF our samples are drawn from populations with the same median, there is a (28%) `r mwu$p.value` probability of finding sample medians at least as different as mine by random chance.

We do not have enough evidence to reject our H0, so we fail to reject. 

There is no significant difference in medians (Mann-Whitney U; U = `r mwu$statistic`, p-value = `r round(mwu$p.value, 2)`).

The medians of group A (M = 22.4 g) and group B (M= 31.2 g) did not differ significantly (Mann Whitney U, ...). 

## Simple linear regression 

Two *continuous quantitative* variables that you are trying to understand the relationship between, and you think that a linear relationship makes sense. 

Use the `penguins` data from {palmerpenguins}, exploring flipper length (x) versus body mass (y).
```{r}
# Use view(penguins) to look at it 
# Make an exploratory scatterplot of penguin flipper length (x) versus body mass (y)
# Once plotted, ask yourself whether a linear relationship makes sense as a way to describe this relationship.

ggplot(data = penguins, aes(x = flipper_length_mm,
                            y = body_mass_g)) +
         geom_point() +
  geom_smooth(method = "lm") 
```
### Find the model estimates (coefficients)

Use the `lm()` function for linear model.

body_mass_g ~ <<<< tilde (not minus sign)
```{r}
penguin_lm <- lm(body_mass_g ~ flipper_length_mm, data = penguins)
# View it in console
# what does it mean? that bodymass (G) = 49.69(FLIPPERLENGTH) - 5780.83
# What must be the units of the 49.69 coefficient for this equation to make sense? The final units for the right side has to be grams.
# 49.69 is that on average, we expect a 1 mm increase on flipper length for every 49.69 increase in body mass grams
# -5780.83 units? grams
# How do you interpret "-5780.83" and why is not valuable at all? that is the mass of the penguin with no flippers. That is the expected average mass of a penguin with flipper length of zero. when the x-axis variable = 0. 
```

Refer anything from the summary to an inline code chunk:
Model coefficients: `r penguin_lm$coefficients`
Flipper length coefficient: `r penguin_lm$coefficients[2]` # isolating one of them

### Use `brom::tidy()` to get a tidy ouput from lm()
```{r}
penguin_lm_tidy <- tidy(penguin_lm)

penguin_lm_tidy
```


### Use broom::glance() to get output of model fit and statistical data.
```{r}
penguin_lm_glance <- glance(penguin_lm)

penguin_lm_glance
```

### Use equationmatic::extract_eq() to return the model in LaTex
LaText is great to write mathematical equations in code.
```{r}
extract_eq(model = penguin_lm, use_coefs = TRUE) # This includes your values into that expression
```

### Use plot() function to automatically produce diagnostic plots 
```{r}
plot(penguin_lm) # homoscedasticity vs heteroscedasity, residual spread
```

